### 2021-08-15 doc/thesis/chapters/deployment-on-hardware.tex
all /ref to ~/ref (to avoid newline separations)

Chapter \ref{cha:Implementation} already described the development of the main IP core, which is the FM radio receiver in the VHDL and HLS variants.
In order to embed these IPs on the actual FPGA, or rather SoC hardware, a significant additional amount of implementation is neccessary.
This section describes these additional parts of the implementation, which specifically enable and support the deployment on the SoC.

@@

The block diagram in Figure \ref{fig:bd_impl_vivado} shows the implemented architecture that is used in the SoC.
The various design domains in and around the SoC are depicted in different visual representations.
The FPGA domain blocks are highlighted in gray, with the main IPs being surrounded by emphasized lines.
The processor system domain is shown as white blocks.
Blocks that are external to the SoC are drawn with dashed lines.

@@

The CPU is the main actor that determines any actions on the system.
First, it reads a file from the external SD card and stores this data in a dedicated area in the memory.
Next, the DMA is configured and its data transfer mode is initiated, so that the DMA starts to transfer data from the mentioned memory area towards the FPGA.
The DMA is configured in a mode that creates a continuous transfer, by looping through the data -- once the end is reached, the transfer starts at the beginning of the memory area.
Any data connection in the following processing chain is implemented as an AXI stream bus.
The first block after the DMA is an AXI Stream Switch.
It can switch its single input to be forwarded to either of the outputs.
The switch is configured by the CPU, to enable one output at a time.
Consequently, one of the FM radio receiver IPs is provided with data.
The IP processes this data to produce an audio output, which is delivered to another AXI Stream Switch.
It fulfills the function to either select the output of above or below IP, in coordination with the first switch, obviously.
The next block is an AXI Stream Broadcaster, which takes a single stream input and simply duplicates it at its output, so that the outputs are an exact copy of the input stream.
One of these duplicates is sourced to an AXI-Stream-To-I2S converter, which generates the I2S signals for the external audio codec chip.
The second duplicate is connected to a FIFO, which simply stores the generated output data.
Once the FIFO is almost full, it asserts an interrupt to the CPU, which triggers it to read out the entire FIFO buffer.
The CPU then stores this data on the external SD card again.
In that way, the loop of data is closed, which enables the verification of the respective IPs' correct functionality.
Therefore, this system design can be described as a hardware-in-the-loop system, in the point of view of the FPGA IPs.
This ability refers to the knowledge gained with the SIMILAR concept, that is introduced in Section \ref{sec:SimilarProcess}.
Specifically, this implemented system design enables the application of the steps 'Assess Performance' and 'Re-evaluate', since the IPs' performance can be measured and their design can be re-evaluated.\\

>>>>

Chapter~\ref{cha:Implementation} already described the development of the main IP core, which is the FM radio receiver in the VHDL and HLS variants.
In order to embed these IPs on the actual FPGA, or rather SoC hardware, a significant additional amount of implementation is neccessary.
This section describes these additional parts of the implementation, which specifically enable and support the deployment on the SoC.

@@

The block diagram in Figure~\ref{fig:bd_impl_vivado} shows the implemented architecture that is used in the SoC.
The various design domains in and around the SoC are depicted in different visual representations.
The FPGA domain blocks are highlighted in gray, with the main IPs being surrounded by emphasized lines.
The processor system domain is shown as white blocks.
Blocks that are external to the SoC are drawn with dashed lines.

@@

The CPU is the main actor that determines any actions on the system.
First, it reads a file from the external SD card and stores this data in a dedicated area in the memory.
Next, the DMA is configured and its data transfer mode is initiated, so that the DMA starts to transfer data from the mentioned memory area towards the FPGA.
The DMA is configured in a mode that creates a continuous transfer, by looping through the data -- once the end is reached, the transfer starts at the beginning of the memory area.
Any data connection in the following processing chain is implemented as an AXI stream bus.
The first block after the DMA is an AXI Stream Switch.
It can switch its single input to be forwarded to either of the outputs.
The switch is configured by the CPU, to enable one output at a time.
Consequently, one of the FM radio receiver IPs is provided with data.
The IP processes this data to produce an audio output, which is delivered to another AXI Stream Switch.
It fulfills the function to either select the output of above or below IP, in coordination with the first switch, obviously.
The next block is an AXI Stream Broadcaster, which takes a single stream input and simply duplicates it at its output, so that the outputs are an exact copy of the input stream.
One of these duplicates is sourced to an AXI-Stream-To-I2S converter, which generates the I2S signals for the external audio codec chip.
The second duplicate is connected to a FIFO, which simply stores the generated output data.
Once the FIFO is almost full, it asserts an interrupt to the CPU, which triggers it to read out the entire FIFO buffer.
The CPU then stores this data on the external SD card again.
In that way, the loop of data is closed, which enables the verification of the respective IPs' correct functionality.
Therefore, this system design can be described as a hardware-in-the-loop system, in the point of view of the FPGA IPs.
This ability refers to the knowledge gained with the SIMILAR concept, that is introduced in Section~\ref{sec:SimilarProcess}.
Specifically, this implemented system design enables the application of the steps 'Assess Performance' and 'Re-evaluate', since the IPs' performance can be measured and their design can be re-evaluated.\\

### 2021-08-15 doc/thesis/chapters/dsp.tex
all /ref to ~/ref (to avoid newline separations)

\noindent
The block diagram in Figure \ref{fig:bd_dsp_overview} gives a high-level overview of a receiver structure.

@@

Figure \ref{fig:bd_superhet} shows a block diagram of a superhet receiver.
The antenna signal first runs through a band selection filter.
It selects the desired frequency band and filters out any unwanted signals outside of this band.
A low-noise amplifier (LNA) may be used to boost the signal strenght.
The next step is the mixer, which is fed by a local oscillator (LO).
This shifts, or down-converts, the antenna signal down to the IF.
Afterwards, the IF signal is amplified and filtered again to remove the side-products of the mixer, namely the image replicas.
Thus, it is also called an image rejection filter.
The demodulator stage can now perform any further signal processing on the signal, until the required signal is retrieved.\\

@@

A block diagram, as well as the signal spectrum of the process is shown in Figure \ref{fig:bd_down_conversion_spectrum}. In this example, the frequency of the LO is set to shift the signal from $f_m$ to baseband.

@@

The procedure that is used for down-conversion is an amplitude modulation (AM).
The equation for a generic AM signal is $ S_{AM}(t) = A_m m(t) \cdot A_c c(t) $, where $m(t)$ represents the message signal, and $c(t)$ is the high-frequency carrier.
Their respective amplitudes are $A_m$ and $A_c$.
This equation is depicted as a block diagram and signal spectrum in Figure \ref{fig:bd_down_conversion_spectrum}, where $S_{AM}$ is denoted as the baseband signal $m_{BB}$ of the original signal $m$.
The signal $m(t)$ is a baseband signal.
It may be antenna signal that was filtered by a bandpass filter.
The received signal is a real signal, and thus, its negative frequency spectrum is an exact mirror of the positive frequencies.
This fact can be explained in two simple equations.\\

@@

The graphical representation in the frequency spectrum is shown in Figure \ref{fig:down_conversion_spectrum_steps}.
As mentioned above, the assumption is a direct conversion to baseband, so $w_m = w_c$.
Looking at the positive original part, it is shown that only half of its energy is translated into the desired baseband.
However, half of the energy of the original negative part is also shifted to baseband.
The superposition of the two parts results in an amount of energy that is again equal to the original signal.

@@

In words: The sampling frequency must be higher than twice the frequency of the signal's highest frequency content.
Any digital signal processing technique must hold the specification of the sampling theorem \cite[Ch. 4.2]{AlessioSilviaMaria2016DSPa}.
In case the theorem is violated, the effect of aliasing occurs on the generated signal.
The sampling process introduces spectral replicas at integer multiples of the sampling frequency $f_s$.
These replicas overlap, if $f_s$ does not meet the theorem, which results in signal distortion.
Figure \ref{fig:aliasing_freq_spectrum} illustrates the effect of aliasing in the frequency spectrum \cite{ThyagarajanK.S2019ItDS}.

@@

The instantanious frequency $f_i$ of an FM signal can be described as
\begin{equation}
  f_i = f_c + \Delta f \cdot m(t)
\end{equation}
where $f_c$ is the carrier- or center frequency, $\Delta f$ is the maximum frequency deviation and $m(t)$ is the information or message signal that is to be transmitted.
Simply looking at this equation, the frequency deviation, which is sometimes also called the swing, varies in the range between $f_c \pm \Delta f$.
Considering the relationship between frequency and angle, as described above, and after some simple substitutions which will not be described into detail here, the equation for a generic frequency modulated wave is the following
\begin{equation}
  s(t) = A_c\ cos \Big( 2 \pi f_c t + 2 \pi k_f \int m(t) dt \Big),
  \label{equ:fm_func}
\end{equation}
where $A_c$ is the amplitude of the resulting FM signal and $k_f$ is the frequency sensitivity factor.
This sensitivity factor has a direct relationship with the modulation index $\beta$.
\begin{equation}
  \beta = \frac{\Delta f}{f_m} = \frac{k_f A_m}{f_m},
\end{equation}
where $f_m$ is the highest existing frequency, or the bandwidth of the information signal.
% More info on modulation index: https://fas.org/man/dod-101/navy/docs/es310/FM.htm
Formula \eqref{equ:fm_func} describes a frequency modulated signal with a generic message signal $m(t)$.
A widely used example application of FM is broadcast radio, where audio streams are transmitted.
An audio stream can be described as a cosine wave.
Therefore, the information signal that is to be transmitted in broadcast radio can be represented as a cosine wave in the form of
\begin{equation}
  m(t) = A_m\ cos(2 \pi f_m t).
\end{equation}
By inserting this message signal into the generic FM Equation \eqref{equ:fm_func}, the final formula for an FM signal transforms into the following form.
\begin{equation}
  y(t) = A_c\ cos \Big(2 \pi f_c t + \beta\ sin(2 \pi f_m t)\Big )
\end{equation}
Several formulas, as well as their derivations are described according to \cite[pg.54-55]{MatinMohammadA2018CSfE}.

@@

Figure \ref{fig:fm_channel_baseband_freqs} shows the frequency spectrum of a single channel, located at its carrier- or center frequency.
For a better overview, only the positive spectum part is shown.
The negative part is an exact mirror, because the transmitted and received signals from the antenna are real signals -- they do not have an imaginary part.
The spectrum consists of multiple parts.
Therefore, it is also referred to as `multiplex signal' in the literature.
Several parts are described in the following paragraphs \cite{EbuFmRegulators},\cite{EtsiFmSoundBroadcastingServiceHarmonizedStandard},\cite{ElectronicsNotesFmTutorial}.

@@

The bandwidth for this difference-signal spans 15~kHz on either side of the subcarrier.
It is used to generate a stereo audio signal, in combination with the mono signal.
This process is explained into more detail in Chapter \ref{sec:demodulation-of-broadcast-fm}.

@@

A frequency discriminator generates an output that is directly proportional to the frequency of the input FM signal.
In other words, it converts FM to AM.
The discriminator consists of two main parts -- a differentiator and an envelope detector.
Different strategies can be implemented for both parts.
A block diagram for a frequency discriminator with its two main parts is shown in Figure \ref{fig:bd_freq_discriminator}.

@@

The differentiator already converts the FM signal into an AM signal, but still leaves an FM content.
The transformation that happens here can be expressed in a formula, by simply differentiating Equation \eqref{equ:fm_func}.
This operation, with some additional simple algebraic transformations, results in
\begin{equation}
  \frac{d s(t)}{dt} = A_c 2 \pi \Big[f_c + k_f m(t) \Big] sin \Big(\omega_c t + 2 \pi k_f \int m(t) dt -pi \Big)
  \label{equ:fm_demod_discriminator}
\end{equation}

@@

A time-domain example for this signal is shown in the top left diagram in Figure \ref{fig:time_domain_envelope_detect}.\\

An important requirement for the differentiator is, that the input signal is of a constant amplitude \cite{SchnyderHaller2002}.
The transmitted signal is subject to random additive noise on the transmission channel.
The differentiator would deliver wrong results because of this noise, when substracting two consecutive samples, as described in Equation \eqref{equ:differentiator_two_samples}.
To achieve a constant amplitude, the received complex baseband signal $s(t)$ can be normalized to a value of one.
\begin{equation}
  s_{norm} = \frac{s}{|s|} = \frac{A(n)\ e^{j\phi_{\text{FM}}(n)}}{|{A(n)\ e^{j\phi_{\text{FM}}(n)}|}} = e^{j\phi_{\text{FM}}(n)}
\end{equation}
\begin{equation}
  |e^{j\phi_{FM}(n)}| = 1
\end{equation}

@@

The differentiated signal still consists of a FM part, as explained above, using Equation \eqref{equ:fm_demod_discriminator}.
In order to remove this high-frequeny part and extract the low-frequency envelope, the signal needs to be fed into an envelope detector.
A block diagram is shown in Figure \ref{fig:bd_freq_discriminator}.\\

The method implemented in Figure \ref{fig:time_domain_envelope_detect} is one of the most simple ones and is usually referred to as `Asynchronous Half-Wave Envelope Detector' \cite{DSPRelatedEnvelopeDetection}.
It consists of a thresholding unit to remove negative values and a conventional lowpass filter.
In analog implementations, the thresholding unit is a diode.
The cutoff frequency of the lowpass needs to be adapted to the envelope signals' maximum frequency.
It needs to be able to follow the signal, but also sufficiently smooth out the rectified signal.\\

@@

This section describes the demodulation of the information content of an FM broadcast channel, as it is used in commercial FM radio broadcasting.
The assumption here is, that the actual FM demodulation, as described in Chapter \ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly, including the de-emphasis of the signal.
So at this point the FM signal is converted to an AM signal.
Thus, traditional AM DSP techniques can be employed to decode the FM broadcast channel.\\

A frequency spectrum of the multiplex signal of an FM channel is shown in Figure \ref{fig:fm_channel_baseband_freqs}.

@@

The block diagram in Figure \ref{fig_bd_stereo_demod} illustrates the signal processing chain for stereo audio.
In the block diagram, $x(t)$ represents the multiplex signal from the FM demodulator.
The signals $x_L(t)$ and $x_R(t)$ describe the left and right audio signal, respectively.\\

@@

In the upper branch, a bandpass filter selects the difference signal with ranges from 23 to 53~kHz.
It is then modulated to baseband by the previously generated 38~kHz subcarrier.
Subsequently, a lowpass limits signal bandwidth to 15~kHz, which removes the generated modulation artifacts.
The lower branch lowpass-filters the summation, or mono signal, as described in Section \ref{subsec:demod_mono}.
The rightmost part in the diagram performs the combination of the mono and the difference signal, according to Equation \eqref{equ_stereo_from_sum_diff}.

>>>>

\noindent
The block diagram in Figure~\ref{fig:bd_dsp_overview} gives a high-level overview of a receiver structure.

@@

Figure~\ref{fig:bd_superhet} shows a block diagram of a superhet receiver.
The antenna signal first runs through a band selection filter.
It selects the desired frequency band and filters out any unwanted signals outside of this band.
A low-noise amplifier (LNA) may be used to boost the signal strenght.
The next step is the mixer, which is fed by a local oscillator (LO).
This shifts, or down-converts, the antenna signal down to the IF.
Afterwards, the IF signal is amplified and filtered again to remove the side-products of the mixer, namely the image replicas.
Thus, it is also called an image rejection filter.
The demodulator stage can now perform any further signal processing on the signal, until the required signal is retrieved.\\

@@

A block diagram, as well as the signal spectrum of the process is shown in Figure~\ref{fig:bd_down_conversion_spectrum}. In this example, the frequency of the LO is set to shift the signal from $f_m$ to baseband.

@@

The procedure that is used for down-conversion is an amplitude modulation (AM).
The equation for a generic AM signal is $ S_{AM}(t) = A_m m(t) \cdot A_c c(t) $, where $m(t)$ represents the message signal, and $c(t)$ is the high-frequency carrier.
Their respective amplitudes are $A_m$ and $A_c$.
This equation is depicted as a block diagram and signal spectrum in Figure~\ref{fig:bd_down_conversion_spectrum}, where $S_{AM}$ is denoted as the baseband signal $m_{BB}$ of the original signal $m$.
The signal $m(t)$ is a baseband signal.
It may be antenna signal that was filtered by a bandpass filter.
The received signal is a real signal, and thus, its negative frequency spectrum is an exact mirror of the positive frequencies.
This fact can be explained in two simple equations.\\

@@

The graphical representation in the frequency spectrum is shown in Figure~\ref{fig:down_conversion_spectrum_steps}.
As mentioned above, the assumption is a direct conversion to baseband, so $w_m = w_c$.
Looking at the positive original part, it is shown that only half of its energy is translated into the desired baseband.
However, half of the energy of the original negative part is also shifted to baseband.
The superposition of the two parts results in an amount of energy that is again equal to the original signal.

@@

In words: The sampling frequency must be higher than twice the frequency of the signal's highest frequency content.
Any digital signal processing technique must hold the specification of the sampling theorem \cite[Ch. 4.2]{AlessioSilviaMaria2016DSPa}.
In case the theorem is violated, the effect of aliasing occurs on the generated signal.
The sampling process introduces spectral replicas at integer multiples of the sampling frequency $f_s$.
These replicas overlap, if $f_s$ does not meet the theorem, which results in signal distortion.
Figure~\ref{fig:aliasing_freq_spectrum} illustrates the effect of aliasing in the frequency spectrum \cite{ThyagarajanK.S2019ItDS}.

@@

The instantanious frequency $f_i$ of an FM signal can be described as
\begin{equation}
  f_i = f_c + \Delta f \cdot m(t)
\end{equation}
where $f_c$ is the carrier- or center frequency, $\Delta f$ is the maximum frequency deviation and $m(t)$ is the information or message signal that is to be transmitted.
Simply looking at this equation, the frequency deviation, which is sometimes also called the swing, varies in the range between $f_c \pm \Delta f$.
Considering the relationship between frequency and angle, as described above, and after some simple substitutions which will not be described into detail here, the equation for a generic frequency modulated wave is the following
\begin{equation}
  s(t) = A_c\ cos \Big( 2 \pi f_c t + 2 \pi k_f \int m(t) dt \Big),
  \label{equ:fm_func}
\end{equation}
where $A_c$ is the amplitude of the resulting FM signal and $k_f$ is the frequency sensitivity factor.
This sensitivity factor has a direct relationship with the modulation index $\beta$.
\begin{equation}
  \beta = \frac{\Delta f}{f_m} = \frac{k_f A_m}{f_m},
\end{equation}
where $f_m$ is the highest existing frequency, or the bandwidth of the information signal.
% More info on modulation index: https://fas.org/man/dod-101/navy/docs/es310/FM.htm
Formula~\eqref{equ:fm_func} describes a frequency modulated signal with a generic message signal $m(t)$.
A widely used example application of FM is broadcast radio, where audio streams are transmitted.
An audio stream can be described as a cosine wave.
Therefore, the information signal that is to be transmitted in broadcast radio can be represented as a cosine wave in the form of
\begin{equation}
  m(t) = A_m\ cos(2 \pi f_m t).
\end{equation}
By inserting this message signal into the generic FM Equation~\eqref{equ:fm_func}, the final formula for an FM signal transforms into the following form.
\begin{equation}
  y(t) = A_c\ cos \Big(2 \pi f_c t + \beta\ sin(2 \pi f_m t)\Big )
\end{equation}
Several formulas, as well as their derivations are described according to \cite[pg.54-55]{MatinMohammadA2018CSfE}.

@@

Figure~\ref{fig:fm_channel_baseband_freqs} shows the frequency spectrum of a single channel, located at its carrier- or center frequency.
For a better overview, only the positive spectum part is shown.
The negative part is an exact mirror, because the transmitted and received signals from the antenna are real signals -- they do not have an imaginary part.
The spectrum consists of multiple parts.
Therefore, it is also referred to as `multiplex signal' in the literature.
Several parts are described in the following paragraphs \cite{EbuFmRegulators},\cite{EtsiFmSoundBroadcastingServiceHarmonizedStandard},\cite{ElectronicsNotesFmTutorial}.

@@

The bandwidth for this difference-signal spans 15~kHz on either side of the subcarrier.
It is used to generate a stereo audio signal, in combination with the mono signal.
This process is explained into more detail in Chapter~\ref{sec:demodulation-of-broadcast-fm}.

@@

A frequency discriminator generates an output that is directly proportional to the frequency of the input FM signal.
In other words, it converts FM to AM.
The discriminator consists of two main parts -- a differentiator and an envelope detector.
Different strategies can be implemented for both parts.
A block diagram for a frequency discriminator with its two main parts is shown in Figure~\ref{fig:bd_freq_discriminator}.

@@

The differentiator already converts the FM signal into an AM signal, but still leaves an FM content.
The transformation that happens here can be expressed in a formula, by simply differentiating Equation~\eqref{equ:fm_func}.
This operation, with some additional simple algebraic transformations, results in
\begin{equation}
  \frac{d s(t)}{dt} = A_c 2 \pi \Big[f_c + k_f m(t) \Big] sin \Big(\omega_c t + 2 \pi k_f \int m(t) dt -pi \Big)
  \label{equ:fm_demod_discriminator}
\end{equation}

@@

A time-domain example for this signal is shown in the top left diagram in Figure~\ref{fig:time_domain_envelope_detect}.\\

An important requirement for the differentiator is, that the input signal is of a constant amplitude \cite{SchnyderHaller2002}.
The transmitted signal is subject to random additive noise on the transmission channel.
The differentiator would deliver wrong results because of this noise, when substracting two consecutive samples, as described in Equation~\eqref{equ:differentiator_two_samples}.
To achieve a constant amplitude, the received complex baseband signal $s(t)$ can be normalized to a value of one.
\begin{equation}
  s_{norm} = \frac{s}{|s|} = \frac{A(n)\ e^{j\phi_{\text{FM}}(n)}}{|{A(n)\ e^{j\phi_{\text{FM}}(n)}|}} = e^{j\phi_{\text{FM}}(n)}
\end{equation}
\begin{equation}
  |e^{j\phi_{FM}(n)}| = 1
\end{equation}

@@

The differentiated signal still consists of a FM part, as explained above, using Equation~\eqref{equ:fm_demod_discriminator}.
In order to remove this high-frequeny part and extract the low-frequency envelope, the signal needs to be fed into an envelope detector.
A block diagram is shown in Figure~\ref{fig:bd_freq_discriminator}.\\

The method implemented in Figure~\ref{fig:time_domain_envelope_detect} is one of the most simple ones and is usually referred to as `Asynchronous Half-Wave Envelope Detector' \cite{DSPRelatedEnvelopeDetection}.
It consists of a thresholding unit to remove negative values and a conventional lowpass filter.
In analog implementations, the thresholding unit is a diode.
The cutoff frequency of the lowpass needs to be adapted to the envelope signals' maximum frequency.
It needs to be able to follow the signal, but also sufficiently smooth out the rectified signal.\\

@@

This section describes the demodulation of the information content of an FM broadcast channel, as it is used in commercial FM radio broadcasting.
The assumption here is, that the actual FM demodulation, as described in Chapter~\ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly, including the de-emphasis of the signal.
So at this point the FM signal is converted to an AM signal.
Thus, traditional AM DSP techniques can be employed to decode the FM broadcast channel.\\

A frequency spectrum of the multiplex signal of an FM channel is shown in Figure~\ref{fig:fm_channel_baseband_freqs}.

@@

The block diagram in Figure~\ref{fig_bd_stereo_demod} illustrates the signal processing chain for stereo audio.
In the block diagram, $x(t)$ represents the multiplex signal from the FM demodulator.
The signals $x_L(t)$ and $x_R(t)$ describe the left and right audio signal, respectively.\\

@@

In the upper branch, a bandpass filter selects the difference signal with ranges from 23 to 53~kHz.
It is then modulated to baseband by the previously generated 38~kHz subcarrier.
Subsequently, a lowpass limits signal bandwidth to 15~kHz, which removes the generated modulation artifacts.
The lower branch lowpass-filters the summation, or mono signal, as described in Section~\ref{subsec:demod_mono}.
The rightmost part in the diagram performs the combination of the mono and the difference signal, according to Equation~\eqref{equ_stereo_from_sum_diff}.

### 2021-08-15 doc/thesis/chapters/hls.tex
all /ref to ~/ref (to avoid newline separations)

A major achievement in the usage of HLS is the shorter development time.
The time that it takes to develop a first working version of the intended design can be significantly reduced, compared to the development in RTL code directly.
This is especially important when an FPGA design is combined with a software part, like in a System On Chip, since software development is comparably fast.
Figure \ref{fig:timeline_performance_first_prototype} schematically displays the difference in development time, over performance and degree of optimization, when the same functionality of a prototype is implemented in different methods for various targets.
The compared targets are an x86 processor CPU, a digital signal processor (DSP), a graphics processing unit (GPU) and the FPGA design in RTL or HLS implementation.
It is clearly visible, that the FPGA design with the HLS design approach achieves a first working prototype much earlier than the direct RTL approach.

@@

The code example in Program \ref{prog:axi4_lite} shows the specification of an AXI4-Lite memory-mapped slave interface.
The provided registers are separated into configuration registers and status registers.
The configuration is readable and writeable, while the status registers are implemented as read-only, which requires less logic in the RTL implementation.\\

@@

The second code example, Program \ref{prog:axi4_stream}, demonstrates the equally simple definition of an AXI4-Stream interface.

@@

Section \ref{sec:hls:workflow} above already mentioned the topics of testbenches and simulation types.
The following descriptions shall be seen as a comprehensive extension around these topics.

>>>>

A major achievement in the usage of HLS is the shorter development time.
The time that it takes to develop a first working version of the intended design can be significantly reduced, compared to the development in RTL code directly.
This is especially important when an FPGA design is combined with a software part, like in a System On Chip, since software development is comparably fast.
Figure~\ref{fig:timeline_performance_first_prototype} schematically displays the difference in development time, over performance and degree of optimization, when the same functionality of a prototype is implemented in different methods for various targets.
The compared targets are an x86 processor CPU, a digital signal processor (DSP), a graphics processing unit (GPU) and the FPGA design in RTL or HLS implementation.
It is clearly visible, that the FPGA design with the HLS design approach achieves a first working prototype much earlier than the direct RTL approach.

@@

The code example in Program~\ref{prog:axi4_lite} shows the specification of an AXI4-Lite memory-mapped slave interface.
The provided registers are separated into configuration registers and status registers.
The configuration is readable and writeable, while the status registers are implemented as read-only, which requires less logic in the RTL implementation.\\

@@

The second code example, Program~\ref{prog:axi4_stream}, demonstrates the equally simple definition of an AXI4-Stream interface.

@@

Section~\ref{sec:hls:workflow} above already mentioned the topics of testbenches and simulation types.
The following descriptions shall be seen as a comprehensive extension around these topics.
